---
title: "Task 3"
author: "Manuela"
date: "24 octobre 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Downloading the packages I need

```{r}
library(readr)
library(tidyverse)
```


##Creating dataframes from the fixed width files

```{r}
US_births2017 <- read_fwf("../../Nat2017PublicUS.c20180516.r20180808.txt",
fwf_positions(start = c(13,475,504),
end = c(14,475,507),
col_names = c("Birth","sex", "weight")))

US_births2018 <-read_fwf("../../Nat2018PublicUS.c20190509.r20190717.txt",
fwf_positions(start = c(13,475,504),
end = c(14,475,507),
col_names = c("Birth","sex", "weight")))
```


#### Proportion of boys to girls in 2017 and 2018

```{r}
Boys2017 <- US_births2017 %>% filter(sex == "M") %>% count()
Girls2017 <- US_births2017 %>% filter(sex == "F") %>% count()
proportionBoys2017 <- Boys2017 / (Boys2017+Girls2017)
proportionGirls2017 <- 1 - proportionBoys2017


Boys2018 <- US_births2018 %>% filter(sex == "M") %>% count()
Girls2018 <- US_births2018 %>% filter(sex == "F") %>% count()
proportionBoys2018 <- Boys2018 / (Boys2018+Girls2018)
proportionGirls2018 <- 1 - proportionBoys2018

```



**The proportion of girls in 2017 was `r proportionGirls2017 * 100` (and `r proportionBoys2017 *100 ` boys ) in 2018 they were `r proportionGirls2018 *100 ` girls (and  `r proportionBoys2018 *100 ` boys). There is no main difference between the number of boys and girls born in 2017 and 2018 **

#### Average birth weight

```{r}
US_births2017$weight <- as.numeric(US_births2017$weight)
US_births2018$weight <- as.numeric(US_births2018$weight)

average_weight2017 <- summarize(US_births2017, average_weight = mean(US_births2017$weight))

average_weight2018 <- summarize(US_births2018, average_weight = mean(US_births2018$weight))
```

**In 2017 the average birth weight was `r average_weight2017`in grams whereas in 2018 it was `r average_weight2018` in grams **


#### Density plot of average births in 2018

```{r}
ggplot(US_births2018, aes(x=weight, fill=sex)) + geom_density() +  ggtitle("US Birth weight of boys and girls in 2018")
```

#### Merging the two dataframes into one to answer the next question more easily

```{r}
US_births2017 <- mutate(US_births2017, year="2017")
US_births2018 <- mutate(US_births2018, year="2018")
Births <- rbind(US_births2017, US_births2018)
```


#### Average birth weigh variation by Month and Year?


#### Mean for month, creating a columnf or month and year 
```{r}
average_weight_m <- data %>% group_by(year, Birth) %>%
summarize(average_weight = mean(weight))%>%
unite("month_year", year, Birth, sep="/")

```

#### Making a line plot to answer the question 

```{r}
ggplot(average_weight_month, aes(x=month_year, y=average_weight)) + geom_line() +
  ggtitle("US average birth weight from 2017 to 2018")
```


**It looks like in the start of the year babies weight less than those who are born in summer**
